---
title: "Education & Career Success"
author: "Tran Dang Huy, Sia Chawla Chawla, Jiaqi Xu"

editor: source
embed-resources: true


# set report format 
format:
  pdf:
    toc: true
    theme: journal
    margin-top: 20mm
    margin-bottom: 25mm
    margin-left: 25mm
    margin-right: 25mm

# code chunk and result settings, globally, (method 1)
execute: 
  echo: false      # hide code chunk
  eval: true       # run the code
  include: true    # show output
  cache: false     
  warning: false
  message: false
  fig-align: "center"   # plot's position: left, center, or right
  fig-width: 6          # plot's width
  fig-height: 4         # plot's height
  out-width: "65%"      # % width of inserted image

# allow sectioning
number-sections: true
section-divs: true
section-depth: 3
---


```{r}
library(knitr)
library(tidyverse)
```


# Executive summary {#sec-summary}

This project examines which student experiences most strongly influence early career success, using a dataset of 5,000 graduates from Kaggle. The analysis identified internships, high school GPA, field of study, networking skill, and university projects completed as the most significant predictors of job offers, while university ranking and soft skills demonstrated limited impact. The random forest model suggests that employers value a combination of practical experience, academic foundation, and selective engagement in co-curricular activities over institutional prestige. These findings challenge conventional assumptions about employability and provide meaningful insights to help students make informed, strategic career decisions.




# Introduction {#sec-intro}
(Maximum 10 sentences)


Graduates today face many questions about how their experiences influence their careers. Grades alone may not determine who gets the best job opportunities or feels happiest at work. This report explores two main questions: Which student experiences are linked to receiving more job offers? And what makes graduates feel satisfied with their careers and achieve a good work-life balance?

To answer these questions, we use data from 5,000 recent graduates. This data includes information about their personal characteristics, grades, internships, certifications, and soft skills, as well as their self-rated career satisfaction and work-life balance. We do not test specific theories but instead look openly for patterns. Our analysis will highlight which experiences seem connected to better career outcomes, as these may help students and educators see how different experiences could affect career paths.



# Methodology {#sec-method}
(Maximum 300 words; Should include a figure and a table and those must be referenced in the text and have adequate captions)


```{r}
data <- read.csv("data/education_career_success.csv")
```


# Results {#sec-result}
(Maximum 200 words. Should include either a figure or a table.)


# Discussion, conclusion and recommendations {#sec-discuss}


# Reference section {#sec-ref}
(Include at least 1 reference.)
 
