---
title: "Identifying the Student Characteristics that Influence Employment Outcomes"
author:
  
  - name: Sia Chawla
    affiliation: Executive Summary, Discussion
    
  - name: Tran Dang Huy
    affiliation: Result Interpretations, References
    
  - name: Jiaqi Xu
    affiliation: Introduction, Methodology, Result Graphics

editor: source
embed-resources: true


# set report format 
format:
  report-pdf:
    toc: true
    theme: journal
    fig-width: 4      
    fig-height: 3    
    margin-top: 20mm
    margin-bottom: 25mm
    margin-left: 25mm
    margin-right: 25mm
    organization: ETC5513 Group Assessment
    

# code chunk and result settings
execute: 
  echo: false      # hide code chunk
  eval: true       # run the code
  include: true    # show output
  cache: false     
  warning: false
  message: false
  fig-align: "center"   # plot's position: left, center, or right
  fig-width: 6          # plot's width
  fig-height: 4         # plot's height
  out-width: "65%"      # % width of inserted image

# allow sectioning
number-sections: true
section-divs: true
section-depth: 3


# allow reference 
bibliography: references.bib
nocite: |
  @xie2025knitr
  @wickham2019tidyverse
  @liaw2002randomforest
---

\newpage

```{r}
library(knitr)
library(tidyverse)
library(randomForest)
library(tidytext)
```

# Executive summary {#sec-summary}

This project examines which student experiences most strongly influence early career success, using a kaggle dataset from @shamim2022education. The analysis identified internships, high school GPA, field of study, networking skill, and university projects completed as the most significant predictors of job offers, while university ranking and soft skills demonstrated limited impact. The random forest model suggests that employers value a combination of practical experience, academic foundation, and selective engagement in co-curricular activities over institutional prestige. These findings challenge conventional assumptions about employability and provide meaningful insights to help students make informed, strategic career decisions.

# Introduction (Jackie) {#sec-intro}

Graduates today face many questions about how their experiences influence their careers. Grades alone may not determine who gets the more job opportunities. This report explores which student experiences are associated with receiving more job offers.

To answer this question, we use a kaggle dataset from @shamim2022education which contains 5,000 records from recent graduates. It includes information about students’ academic background, personal demographic, and career-related outcomes. Rather than testing predefined theories, this project takes an open-ended, pattern-oriented approach. The aim is to explore which types of experiences appear most consistently linked to job outcomes and personal satisfaction.

# Methodology {#sec-method}

## Data exploration {#sec-dataExplore}

```{r}
data <- read.csv("data/education_career_success.csv")
```

```{r}
#| eval: false
sapply(data, n_distinct)
```

The data was collected from [Kaggle](https://www.kaggle.com/datasets/adilshamim8/education-and-career-success) and initially contains records for `r nrow(data)` recent university graduates. During our data processing, we found 3.8% (`r sum(data$Gender == "Other")` rows) of the data have inconsistent gender values, as shown in @fig-genderDistribution. Since this is likely to provide misleading information, we decided to exclude them from further analysis.

```{r}
#| label: fig-genderDistribution
#| fig-cap: "Number of Gender Categories"

data %>%
  count(Gender) %>%
  mutate(perc = round(100 * n / sum(n), 1),
         label = paste0(Gender, ": ", perc, "%")) %>%
ggplot(aes(x = "", y = n, fill = Gender)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5)) +
  labs(x = NULL, y = NULL) +
  theme_void() +
  theme(legend.position = "none")
```

```{r}
# filter gender variable, and...
# convert text liked variable into categorical variable
data <- data %>%
  dplyr::filter(Gender %in% c("Male", "Female") ) %>%
  mutate(across(where(is.character), as.factor))

# convert ordinal variable to ordered factor 
data <- data %>%
  mutate(
    Career_Satisfaction = factor(Career_Satisfaction, ordered = TRUE),
    Soft_Skills_Score = factor(Soft_Skills_Score, ordered = TRUE),
    Networking_Score = factor(Networking_Score, ordered = TRUE),
    Current_Job_Level = factor(Current_Job_Level,
                               levels = c("Entry", "Mid", "Senior", "Executive"), ordered = TRUE),
    Work_Life_Balance = factor(Work_Life_Balance, order = TRUE),
  )
```

## Model selection {#sec-dataTypeDataModel}

@tbl-data presents the first five records in the dataset and highlights the diversity of input variable types. The dataset includes numerical variables such as `SAT_Score`, categorical fields like `Field_of_Study`, and ordinal variables such as `Networking_Score`. This variety makes it unsuitable for models that require specific data type.

```{r}
#| label: tbl-data
#| tbl-cap: "First 5 records in filtered dataset"

df <- as.data.frame(t(head(data, 5)))
knitr::kable(df)
```

@fig-outcomeskew shows the distribution of the outcome variable, `Job_Offers`, which represents the number of job offers received. The distribution is relatively balanced across classes, meaning no class dominates the data, and no transformation is required. Furthermore, since `Job_Offers` is ordinal, it emphasises the use of classification model over regression model.

```{r}
#| label: fig-outcomeskew
#| fig-cap: "Distributions of Outcome Variable"

data %>%
  dplyr::select(Job_Offers) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  ggplot(aes(x = Value)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
  theme_minimal() +
  labs(x = "Number of Job Offers", y = "Number of Students")
```

Given these characteristics, random forest is an appropriate modelling choice. It is a flexible and robust algorithm that can naturally handle both numerical and categorical inputs without extensive preprocessing. In addition, it captures non-linear relationships and performs well even when input variables vary in scale or measurement type.

## Accesing associations {#sec-importance}

After fitting the random forest model, we use permutation importance to rank all input variables (@fig-rfimportance), as referred by @gregorutti2017correlation. This method measures how much the model’s accuracy drops when the values of one variable are randomly shuffled. A larger drop means the variable was more important for the model’s predictions. This allows us to assess which factors the model relies on most when estimating job offers.

The top 5 predictors were selected based on their importance scores (@fig-rfimportance) and further explored using boxplot with trend lines for the direction of association (@fig-trend).

# Results {#sec-result}

## Associated factors for receiving job offers. {#sec-rfForImportantFactors}

```{r}

outcome_var <- "Job_Offers"

predictors <- data %>%
  dplyr::select(-Student_ID, -Career_Satisfaction, -Work_Life_Balance) %>%
  dplyr::select(-all_of(outcome_var)) %>%
  names()

form <- as.formula(paste(outcome_var, "~", paste(predictors, collapse = " + ")))

set.seed(3)
rf_model <- randomForest(form, data = data, importance = TRUE)

```

@fig-rfimportance ranks predictors by their importance in the random forest model. Students’ practical experiences (Completed Internships and Academic Projects), academic background (High School GPA and Field of Study), and interpersonal networking skills are more associated with the number of job offers received. In contrast, factors such as University Ranking and Soft Skills Score appear to have minimal influence.

```{r}
#| label: fig-rfimportance
#| fig-cap: "Importance of Predictors for Number of Job Offers Received"

rf_imp_df <- importance(rf_model, type = 1) %>%
  as.data.frame() %>%
  rownames_to_column("Variable") %>%
  rename(Importance = `%IncMSE`) %>%  
  arrange(desc(across(where(is.numeric))))

ggplot(rf_imp_df, aes(x = reorder(Variable, Importance), y = Importance)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Predictor",
    y = "% Increase in Accuracy"
  )
```

## Association between important factors and job offers. {#sec-visualiseTrend}

@fig-boxplot and @fig-mean further illustrate how job offers vary across the top five predictors.

```{r}
data <- data %>%
  mutate(HS_GPA_Level = cut(
    High_School_GPA,
    breaks = c(0, 0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4),
    labels = c( "0-0.5", "0.5-1", "1-1.5", "1.5-2", "2-2.5", "2.5-3", "3-3.5", "3.5-4"),
    right = TRUE,
    ordered_result = TRUE
  ))

data_plot <- data %>%
  select(Job_Offers, Internships_Completed, HS_GPA_Level, Field_of_Study, Networking_Score, Projects_Completed) %>%
  mutate(across(-Job_Offers, as.character)) %>%    
  pivot_longer(-Job_Offers, names_to = "Variable", values_to = "Category")


gpa_levels <- c("1.5-2", "2-2.5", "2.5-3", "3-3.5", "3.5-4")

data_plot_sorted <- data_plot %>%
  group_by(Variable) %>%
  group_modify(~{
    facet_df <- .x               # rows for this facet
    facet_nm <- .y$Variable[1]   # facet name (group key)

    if (facet_nm == "HS_GPA_Level") {
      # apply manual order
      facet_df %>%
        mutate(Category = factor(Category, levels = gpa_levels))
    } else {
      # automatic rule: numeric facets in ascending order,
      # text facets keep original order
      Category_num <- suppressWarnings(as.numeric(facet_df$Category))

      level_order <- if (all(is.na(Category_num))) {
        unique(facet_df$Category)                             # text → keep entry order
      } else {
        unique(facet_df$Category[order(Category_num, na.last = NA)])  # numbers → asc
      }

      facet_df %>%
        mutate(Category = factor(Category, levels = level_order))
    }
  }) %>%
  ungroup()
```

The boxplot (@fig-boxplot) shows that the median number of job offers falls between 2 and 3 for most categories. The plot also reveals a downward trend in medians as networking scores and project counts increase.

```{r}
#| label: fig-boxplot
#| fig-cap: "Distribution of job offers across student categories"
#| fig-width: 8   
#| fig-height: 3


data_plot_sorted %>%
  ggplot(aes(x = Category, y = Job_Offers)) +
  geom_boxplot() +
  facet_wrap(~ Variable, scales = "free_x", nrow = 1,
             labeller = as_labeller(c(
                "Field_of_Study" = "Field of Study",
                "HS_GPA_Level" = "High School GPA",
                "Internships_Completed" = "Internships Completed",
                "Networking_Score" = "Networking Skill",
                "Projects_Completed" = "Uni Projects Completed"
              ))) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = NULL, y = "Number of Job Offers") 
  
```

While the patterns for high school GPA and internship experience are more subtle in the boxplot, @fig-mean shows clear upward trends, suggesting that students with higher GPAs and more internships receive more job offers on average.

Among fields of study, Medicine leads with the highest average job offers, followed by Arts and Mathematics, while Computer Science ranks the lowest.

```{r}
#| label: fig-mean
#| fig-cap: "Average number of job offers by predictor category"
#| fig-width: 8   
#| fig-height: 3


data_plot_sorted %>%
  group_by(Variable, Category) %>%
  summarise(avg_offer = mean(Job_Offers), .groups = "drop") %>%
  group_by(Variable) %>%
  mutate(
    line_ok = Variable != "Field_of_Study",
    Category = {
      raw_cat <- as.character(Category)
      if (first(Variable) == "HS_GPA_Level") {
        left_val <- as.numeric(sub("-.*", "", raw_cat))
        factor(Category, levels = raw_cat[order(left_val)])
      } else if (all(grepl("^[0-9.]+$", raw_cat))) {
        factor(Category, levels = raw_cat[order(as.numeric(raw_cat))])
      } else if (all(grepl("^[0-9.]+-[0-9.]+$", raw_cat))) {
        right_val <- as.numeric(sub(".*-", "", raw_cat))
        factor(Category, levels = raw_cat[order(-right_val)])
      } else {
        factor(Category, levels = raw_cat[order(avg_offer, decreasing = TRUE)])
      }
    }
  ) %>%
  ungroup() %>%
  ggplot(aes(x = Category, y = avg_offer, group = Variable)) +
  geom_point(color = "steelblue", size = 2.5) +
  geom_line(data = function(df) dplyr::filter(df, line_ok), aes(group = 1), color = "red") +
  facet_wrap(~ Variable, scales = "free_x", nrow = 1,
             labeller = as_labeller(c(
               "Field_of_Study" = "Field of Study",
               "HS_GPA_Level" = "High School GPA",
               "Internships_Completed" = "Internships Completed",
               "Networking_Score" = "Networking Skill",
               "Projects_Completed" = "Uni Projects Completed"
             ))) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = NULL,
    y = "Average Number of Job Offers"
  )
```

# Discussion, conclusion and recommendations {#sec-discuss}

## Discussion and Conclusion

The analysis revealed that internships completed, high school GPA, field of study, networking skill, and university projects completed were the most influential predictors of job offers. Internships and a strong academic foundation, particularly in high school, were consistently associated with better outcomes, indicating that employers value both preparedness and discipline. While field of study also shaped job offer patterns, the effect was relatively stable across disciplines. Interestingly, both networking skill and number of university projects showed a downward trend beyond moderate levels, suggesting that overinvestment in these areas may yield diminishing returns. These results highlight the nuanced ways in which various student experiences contribute to employability.

In conclusion, early career success is most strongly driven by a balance of practical experience, academic performance, and strategic engagement in co-curricular activities.

## Recommendations

-   **Students** are encouraged to pursue internships and maintain strong academic performance, particularly during high school, while engaging strategically in networking and project-based activities to avoid diminishing returns from overextension.

-   **Higher education institutions** should embed structured experiential learning—such as industry placements and capstone projects—into academic programs, and ensure alignment between fields of study and evolving labour market demands.

-   **Career development services** should provide targeted support that promotes a balanced portfolio of academic achievement, practical experience, and purposeful co-curricular involvement to optimise graduate employability.


# Reference {#sec-ref}

