---
title: "Education & Career Success"
author: "Tran Dang Huy, Sia Chawla Chawla, Jiaqi Xu"

editor: source
embed-resources: true


# set report format 
format:
  pdf:
    toc: true
    theme: journal
    margin-top: 20mm
    margin-bottom: 25mm
    margin-left: 25mm
    margin-right: 25mm

# code chunk and result settings, globally, (method 1)
execute: 
  echo: false      # hide code chunk
  eval: true       # run the code
  include: true    # show output
  cache: false     
  warning: false
  message: false
  fig-align: "center"   # plot's position: left, center, or right
  fig-width: 6          # plot's width
  fig-height: 4         # plot's height
  out-width: "65%"      # % width of inserted image

# allow sectioning
number-sections: true
section-divs: true
section-depth: 3
---

```{r}
library(knitr)
library(tidyverse)
```

# Executive summary {#sec-summary}

(Maximum of 4 sentences)

# Introduction {#sec-intro}

(Maximum 10 sentences)

Graduates today face many questions about how their experiences influence their careers. Grades alone may not determine who gets the best job opportunities or feels happiest at work. This report explores two main questions: Which student experiences are linked to receiving more job offers? And what makes graduates feel satisfied with their careers and achieve a good work-life balance?

To answer these questions, we use data from 5,000 recent graduates. This data includes information about their personal characteristics, grades, internships, certifications, and soft skills, as well as their self-rated career satisfaction and work-life balance. We do not test specific theories but instead look openly for patterns. Our analysis will highlight which experiences seem connected to better career outcomes, as these may help students and educators see how different experiences could affect career paths.

# Methodology {#sec-method}

(Maximum 300 words; Should include a figure and a table and those must be referenced in the text and have adequate captions)

```{r}
data <- read.csv("data/education_career_success.csv")
```

# Results {#sec-result}

(Maximum 200 words. Should include either a figure or a table.)

# Discussion, conclusion and recommendations {#sec-discuss}

**Discussion and Conclusion**

The analysis revealed that internships, high school GPA, and field of study were the most influential predictors of job offers, indicating a shift in employer focus toward practical experience and core academic performance. Internships enable students to apply their knowledge in real-world settings, while high school achievement may signal foundational discipline and capability. In contrast, soft skills and university prestige, though valued, appear less directly tied to immediate employability. The model reinforces the view that tangible experience and academic readiness carry more weight than reputational indicators in early career outcomes.

In conclusion, applied experience and foundational academic strength play a more decisive role in graduate success than institutional status or soft skill measures.

**Recommendations**

-   **Students** are encouraged to actively pursue internships and maintain strong academic performance, particularly during high school, to enhance their employability and demonstrate readiness to prospective employers.


-   **Higher education institutions** should integrate structured experiential learning opportunities—such as industry placements, capstone projects, and applied coursework—while also providing clearer academic pathways aligned with in-demand fields of study.

-   **Career development services** should offer tailored guidance that emphasises the importance of both practical experience and foundational academic performance, equipping graduates to compete more effectively in the job market.

# Reference section {#sec-ref}

(Include at least 1 reference.)
