---
title: "Education & Career Success"
author: "Tran Dang Huy, Sia Chawla Chawla, Jiaqi Xu"

editor: source
embed-resources: true


# set report format 
format:
  pdf:
    toc: true
    theme: journal
    margin-top: 20mm
    margin-bottom: 25mm
    margin-left: 25mm
    margin-right: 25mm

# code chunk and result settings, globally, (method 1)
execute: 
  echo: false      # hide code chunk
  eval: true       # run the code
  include: true    # show output
  cache: false     
  warning: false
  message: false
  fig-align: "center"   # plot's position: left, center, or right
  fig-width: 6          # plot's width
  fig-height: 4         # plot's height
  out-width: "65%"      # % width of inserted image

# allow sectioning
number-sections: true
section-divs: true
section-depth: 3
---


```{r}
library(knitr)
library(tidyverse)
```


# Executive summary {#sec-summary}
(Maximum of 4 sentences)


# Introduction {#sec-intro}
(Maximum 10 sentences)


# Methodology {#sec-method}
(Maximum 300 words; Should include a figure and a table and those must be referenced in the text and have adequate captions)


```{r}
data <- read.csv("data/education_career_success.csv")
```


# Results {#sec-result}
(Maximum 200 words. Should include either a figure or a table.)


# Discussion, conclusion and recommendations {#sec-discuss}


# Reference section {#sec-ref}
(Include at least 1 reference.)
 
