---
title: "Education & Career Success"
author: "Tran Dang Huy, Sia Chawla Chawla, Jiaqi Xu"

editor: source
embed-resources: true


# set report format 
format:
  pdf:
    toc: true
    theme: journal
    margin-top: 20mm
    margin-bottom: 25mm
    margin-left: 25mm
    margin-right: 25mm

# code chunk and result settings, globally, (method 1)
execute: 
  echo: false      # hide code chunk
  eval: true       # run the code
  include: true    # show output
  cache: false     
  warning: false
  message: false
  fig-align: "center"   # plot's position: left, center, or right
  fig-width: 6          # plot's width
  fig-height: 4         # plot's height
  out-width: "65%"      # % width of inserted image

# allow sectioning
number-sections: true
section-divs: true
section-depth: 3
---

```{r}
library(knitr)
library(tidyverse)
```

# Executive summary {#sec-summary}

(Maximum of 4 sentences)

# Introduction {#sec-intro}

(Maximum 10 sentences)

Graduates today face many questions about how their experiences influence their careers. Grades alone may not determine who gets the best job opportunities or feels happiest at work. This report explores two main questions: Which student experiences are linked to receiving more job offers? And what makes graduates feel satisfied with their careers and achieve a good work-life balance?

To answer these questions, we use data from 5,000 recent graduates. This data includes information about their personal characteristics, grades, internships, certifications, and soft skills, as well as their self-rated career satisfaction and work-life balance. We do not test specific theories but instead look openly for patterns. Our analysis will highlight which experiences seem connected to better career outcomes, as these may help students and educators see how different experiences could affect career paths.

# Methodology {#sec-method}

(Maximum 300 words; Should include a figure and a table and those must be referenced in the text and have adequate captions)

```{r}
data <- read.csv("data/education_career_success.csv")
```

# Results {#sec-result}

(Maximum 200 words. Should include either a figure or a table.)

# Discussion, conclusion and recommendations {#sec-discuss}

**Discussion and Conclusion**

The analysis revealed that internships completed, high school GPA, field of study, networking skill, and university projects completed were the most influential predictors of job offers. Internships and a strong academic foundation, particularly in high school, were consistently associated with better outcomes, indicating that employers value both preparedness and discipline. While field of study also shaped job offer patterns, the effect was relatively stable across disciplines. Interestingly, both networking skill and number of university projects showed a downward trend beyond moderate levels, suggesting that overinvestment in these areas may yield diminishing returns. These results highlight the nuanced ways in which various student experiences contribute to employability.

In conclusion, early career success is most strongly driven by a balance of practical experience, academic performance, and strategic engagement in co-curricular activities.

**Recommendations**

-   **Students** are encouraged to pursue internships and maintain strong academic performance, particularly during high school, while engaging strategically in networking and project-based activities to avoid diminishing returns from overextension.


-   **Higher education institutions** sshould embed structured experiential learning—such as industry placements and capstone projects—into academic programs, and ensure alignment between fields of study and evolving labour market demands.

-   **Career development services** should provide targeted support that promotes a balanced portfolio of academic achievement, practical experience, and purposeful co-curricular involvement to optimise graduate employability.

# Reference section {#sec-ref}

(Include at least 1 reference.)
