---
title: "What Associates With More Job Offers?"
subtitle: "Exploring University Student Behaviours and Attributes"
author: "Sia Chawla Chawla,Tran Dang Huy, Jiaqi Xu"
format: 
  revealjs:
    toc: true
    smaller: true
    scrollable: true
    multiplex: false
    show-slide-number: all
    show-notes: false
    controls: true
    width: 1280
    height: 750

    
embed-resources: true


# code chunk and result settings
execute: 
  echo: false      # hide code chunk
  eval: true       # run the code
  include: true    # show output
  cache: false     
  warning: false
  message: false
---

## Introduction

```{r}
library(tidyverse)
```


::: {.callout-note title="Graduate Employment in Australia (2023-2024)"}

According to [ABS](https://www.abs.gov.au/statistics/people/education/education-and-work-australia/latest-release) and [Ministers of the Education Portfolio](https://ministers.education.gov.au/clare/graduate-employment-rates-hit-highest-level-more-decade):

- **84%** of Australians with a `non-school qualification` in 2023 were employed in 2024.
- **78.5%** of `domestic graduates` and **57.7%** of `international graduates` found full-time job 4-6 months after graduation 
:::

:::{.fragment}
These high employment rates suggest that **academic grades alone may no longer be the key differentiator** in job outcomes.  
:::



:::{.callout-tip title = "Research Question" .fragment}
**What student behaviours, attributes, or practices are most associated with receiving more job offers?** 
:::

---

## Dataset Description 

::: {.callout-note title="Data Source"}
Kaggle - [Education and Career Success](https://www.kaggle.com/datasets/adilshamim8/education-and-career-success)  
:::

:::{.columns}

:::{.column width="65%"}

```{r}
data = read.csv("data/education_career_success.csv")
glimpse(data)
```
::: 

:::{.column width="35%"}

Dataset containing categorical, ordinal, and continuous variables.

- Demographic
- Academic Performance
- Skills and Experiences
- Career Outcome

The outcome, `Job_Offers`, is a counting variable

:::

:::

--- 

## Methodology

::: {.columns}

::: {.column width="40%"}

**Data Exploration**:

- Rows with `Gender = "Other"` (3.8%) are excluded due to small sample size  

**Variable Importance**:

- `Random forest` with `permutation`  
  - Handles mixed data types  
  - Captures non-linear relationships  

**Direction of Association**: 

- `Boxplot` and `scatterplot`
  - Suitable for count data  
  - Show group-wise patterns  

:::

::: {.column width="30%"}
```{r}
#| label: fig-genderDistribution
#| fig-cap: "Distribution of Gender"
#| fig-width: 4
#| fig-align: center

data %>%
  count(Gender) %>%
  mutate(
    perc = round(100 * n / sum(n), 1),
    label = paste0(Gender, ": ", perc, "%")
  ) %>%
  ggplot(aes(x = "", y = n, fill = Gender)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  geom_text(aes(label = label), 
            position = position_stack(vjust = 0.5), 
            fontface = "bold", 
            size = 5) +
  labs(x = NULL, y = NULL) +
  theme_void() +
  theme(
    legend.position = "none",
    plot.margin = margin(t = 0, r = 0, b = 0, l = 0)
  )
```

:::

::: {.column width="30%"}
```{r}
#| label: fig-outcomeskew
#| fig-cap: "Distributions of Outcome Variable"
#| fig-width: 4
#| fig-align: left


data %>%
  dplyr::select(Job_Offers) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  ggplot(aes(x = Value)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
  theme_minimal() +
  labs(x = "Number of Job Offer Received", y = "Student Counts", title = NULL)
```
:::

:::











<!-- :::{.fragment} -->
<!-- ```{r} -->
<!-- #| label: fig-genderDistribution -->
<!-- #| fig-cap: "Distribution of Gender" -->
<!-- #| fig-width: 4 -->
<!-- #| fig-height: 4 -->

<!-- data %>% -->
<!--   count(Gender) %>% -->
<!--   mutate( -->
<!--     perc = round(100 * n / sum(n), 1), -->
<!--     label = paste0(Gender, ": ", perc, "%") -->
<!--   ) %>% -->
<!--   ggplot(aes(x = "", y = n, fill = Gender)) + -->
<!--   geom_col(width = 1, color = "white") + -->
<!--   coord_polar(theta = "y") + -->
<!--   geom_text(aes(label = label),  -->
<!--             position = position_stack(vjust = 0.5),  -->
<!--             fontface = "bold",  -->
<!--             size = 5) + -->
<!--   labs(x = NULL, y = NULL) + -->
<!--   theme_void() + -->
<!--   theme( -->
<!--     legend.position = "none", -->
<!--     plot.margin = margin(t = 0, r = 0, b = 0, l = 0) -->
<!--   ) -->
<!-- ``` -->
<!-- ::: -->

<!-- :::{.fragment} -->
<!-- Records with inconsistent value for `gender` were removed.  -->
<!-- ::: -->

<!-- ::: -->
<!-- ::: -->











<!-- ::: {.columns} -->

<!-- ::: {.column width="30%"} -->


<!-- ::: -->

<!-- ::: {.column width="50%"} -->
<!-- #### Row 2, Col 2 -->
<!-- More content here -->
<!-- ::: -->

<!-- ::: -->

---

## page 6

---

## page 7

---

## page 8

---

## Conclusion

::: {.fragment}
- Practical experience and academic performance drive job offers.
:::

::: {.fragment}
- Internships and high school GPA are strong predictors.
:::

::: {.fragment}
- Field of study, networking, and projects matter—if approached strategically.
:::

::: {.fragment}
- Institutional prestige and soft skills had limited impact.
:::


---

## Recommendations

::: {.fragment}
- **Students** must focus on internships, GPA, and **balanced involvement** in co-curriculars.
:::

::: {.fragment}
- **Universities** should embed industry-linked experiences and align programs with job market needs.
:::

::: {.fragment}
- **Career Services** must promote strategic engagement beyond academics.
:::
