---
title: "What Associates With More Job Offers?"
subtitle: "Exploring University Student Behaviours and Attributes"
author: "Sia Chawla, Tran Dang Huy (Andy), Jiaqi Xu (Jackie)"
format: 
  revealjs:
    toc: true
    smaller: true
    scrollable: true
    multiplex: false
    show-slide-number: all
    show-notes: false
    controls: true
    width: 1280
    height: 750

    
embed-resources: true


# code chunk and result settings
execute: 
  echo: false      # hide code chunk
  eval: true       # run the code
  include: true    # show output
  cache: false     
  warning: false
  message: false
---

## Introduction

```{r}
library(tidyverse)
library(knitr)
library(tidyverse)
library(randomForest)
```


::: {.callout-note title="Graduate Employment in Australia (2023-2024)"}

According to [ABS](https://www.abs.gov.au/statistics/people/education/education-and-work-australia/latest-release) and [Ministers of the Education Portfolio](https://ministers.education.gov.au/clare/graduate-employment-rates-hit-highest-level-more-decade):

- **84%** of Australians with a `non-school qualification` in 2023 were employed in 2024.
- **78.5%** of `domestic graduates` and **57.7%** of `international graduates` found full-time job 4-6 months after graduation 
:::

:::{.fragment}
These high employment rates suggest that **academic grades alone may no longer be the key differentiator** in job outcomes.  
:::



:::{.callout-tip title = "Research Question" .fragment}
**What student behaviours, attributes, or practices are most associated with receiving more job offers?** 
:::

---

## Dataset Description 

::: {.callout-note title="Data Source"}
Kaggle - [Education and Career Success](https://www.kaggle.com/datasets/adilshamim8/education-and-career-success)  
:::

:::{.columns}

:::{.column width="65%"}

```{r}
data = read.csv("data/education_career_success.csv")
glimpse(data)
```
::: 

:::{.column width="35%"}

This dataset contains categorical, discrete count, and continuous variables.

:::{.incremental}
- Demographic
- Academic Performance
- Skills and Experiences
- Career Outcome
:::
:::{.fragment}
The outcome, `Job_Offers`, is a counting variable
:::
:::

:::

--- 


## Methodology

::: {.columns}

::: {.column width="40%"}

:::{.fragment}
**Data Exploration**:

- Rows with `Gender = "Other"` (3.8%) are excluded due to small sample size  
:::

:::{.fragment}
**Variable Importance**:

- `Random forest` with `permutation`  
  - Handles mixed data types  
  - Captures not just linear relationships  
:::

:::{.fragment}
**Direction of Association**: 

- `Boxplot` and `scatterplot`
  - Suitable for count data  
  - Show group-wise patterns  
:::
:::

::: {.column width="30%"}
```{r}
#| label: fig-genderDistribution
#| fig-cap: "Distribution of Gender"
#| fig-width: 4
#| fig-align: center

data %>%
  count(Gender) %>%
  mutate(
    perc = round(100 * n / sum(n), 1),
    label = paste0(Gender, ": ", perc, "%")
  ) %>%
  ggplot(aes(x = "", y = n, fill = Gender)) +
  geom_col(width = 1, color = "white") +
  coord_polar(theta = "y") +
  geom_text(aes(label = label), 
            position = position_stack(vjust = 0.5), 
            fontface = "bold", 
            size = 5) +
  labs(x = NULL, y = NULL) +
  theme_void() +
  theme(
    legend.position = "none",
  )
```

:::

::: {.column width="30%"}
```{r}
#| label: fig-outcomeskew
#| fig-cap: "Distributions of Outcome Variable"
#| fig-width: 4
#| fig-align: left


data %>%
  dplyr::select(Job_Offers) %>%
  pivot_longer(cols = everything(), names_to = "Variable", values_to = "Value") %>%
  ggplot(aes(x = Value)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
  theme_minimal() +
  labs(x = "Number of Job Offer Received", y = "Student Counts", title = NULL)
```
:::

:::
---

## Elements that contribute to receiving job offers {.smaller}

::::: columns
::: {.column width="60%"}
```{r, echo=FALSE}

data <- data %>%
  dplyr::filter(Gender %in% c("Male", "Female") ) %>%
  mutate(across(where(is.character), as.factor))

data <- data %>%
  mutate(
    Career_Satisfaction = factor(Career_Satisfaction, ordered = TRUE),
    Soft_Skills_Score = factor(Soft_Skills_Score, ordered = TRUE),
    Networking_Score = factor(Networking_Score, ordered = TRUE),
    Current_Job_Level = factor(Current_Job_Level,
                               levels = c("Entry", "Mid", "Senior", "Executive"), ordered = TRUE),
    Work_Life_Balance = factor(Work_Life_Balance, order = TRUE),
  )

outcome_var <- "Job_Offers"

predictors <- data %>%
  dplyr::select(-Student_ID, -Career_Satisfaction, -Work_Life_Balance) %>%
  dplyr::select(-all_of(outcome_var)) %>%
  names()

form <- as.formula(paste(outcome_var, "~", paste(predictors, collapse = " + ")))

set.seed(3)
rf_model <- randomForest(form, data = data, importance = TRUE)
rf_imp_df <- importance(rf_model, type = 1) %>%
  as.data.frame() %>%
  rownames_to_column("Variable") %>%
  rename(Importance = `%IncMSE`) %>%  
  arrange(desc(across(where(is.numeric))))

ggplot(rf_imp_df, aes(x = reorder(Variable, Importance), y = Importance)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Predictor",
    y = "% Increase in Accuracy"
  )
```
:::

::: {.column width="40%"}
-   This plot shows the importance of each predictor variable in the **random forest model** by using permutation.
-   The result shows that most important factors are related to your **academic performance** and **practical experience**.
-   It is also interesting to noted that **networking skill** rank much higher than **soft skills**.
:::
:::::

---

## Association between important factors and job offers {.smaller}

::::: columns
::: {.column width="50%"}
```{r, echo=FALSE}
data <- data %>%
  mutate(HS_GPA_Level = cut(
    High_School_GPA,
    breaks = c(0, 0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4),
    labels = c( "0-0.5", "0.5-1", "1-1.5", "1.5-2", "2-2.5", "2.5-3", "3-3.5", "3.5-4"),
    right = TRUE,
    ordered_result = TRUE
  ))

data_plot <- data %>%
  select(Job_Offers, Internships_Completed, HS_GPA_Level, Field_of_Study, Networking_Score, Projects_Completed) %>%
  mutate(across(-Job_Offers, as.character)) %>%    
  pivot_longer(-Job_Offers, names_to = "Variable", values_to = "Category")


gpa_levels <- c("1.5-2", "2-2.5", "2.5-3", "3-3.5", "3.5-4")

data_plot_sorted <- data_plot %>%
  group_by(Variable) %>%
  group_modify(~{
    facet_df <- .x               # rows for this facet
    facet_nm <- .y$Variable[1]   # facet name (group key)

    if (facet_nm == "HS_GPA_Level") {
      # apply manual order
      facet_df %>%
        mutate(Category = factor(Category, levels = gpa_levels))
    } else {
      # automatic rule: numeric facets in ascending order,
      # text facets keep original order
      Category_num <- suppressWarnings(as.numeric(facet_df$Category))

      level_order <- if (all(is.na(Category_num))) {
        unique(facet_df$Category)                             # text → keep entry order
      } else {
        unique(facet_df$Category[order(Category_num, na.last = NA)])  # numbers → asc
      }

      facet_df %>%
        mutate(Category = factor(Category, levels = level_order))
    }
  }) %>%
  ungroup()
data_plot_sorted %>%
  ggplot(aes(x = Category, y = Job_Offers)) +
  geom_boxplot() +
  facet_wrap(~ Variable, scales = "free_x", nrow = 1,
             labeller = as_labeller(c(
                "Field_of_Study" = "Field of Study",
                "HS_GPA_Level" = "High School GPA",
                "Internships_Completed" = "Internships Completed",
                "Networking_Score" = "Networking Skill",
                "Projects_Completed" = "Uni Projects Completed"
              ))) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = NULL, y = "Number of Job Offers") 
```
:::

::: {.column width="50%"}
```{r, echo=FALSE}
data_plot_sorted %>%
  group_by(Variable, Category) %>%
  summarise(avg_offer = mean(Job_Offers), .groups = "drop") %>%
  group_by(Variable) %>%
  mutate(
    line_ok = Variable != "Field_of_Study",
    Category = {
      raw_cat <- as.character(Category)
      if (first(Variable) == "HS_GPA_Level") {
        left_val <- as.numeric(sub("-.*", "", raw_cat))
        factor(Category, levels = raw_cat[order(left_val)])
      } else if (all(grepl("^[0-9.]+$", raw_cat))) {
        factor(Category, levels = raw_cat[order(as.numeric(raw_cat))])
      } else if (all(grepl("^[0-9.]+-[0-9.]+$", raw_cat))) {
        right_val <- as.numeric(sub(".*-", "", raw_cat))
        factor(Category, levels = raw_cat[order(-right_val)])
      } else {
        factor(Category, levels = raw_cat[order(avg_offer, decreasing = TRUE)])
      }
    }
  ) %>%
  ungroup() %>%
  ggplot(aes(x = Category, y = avg_offer, group = Variable)) +
  geom_point(color = "steelblue", size = 2.5) +
  geom_line(data = function(df) dplyr::filter(df, line_ok), aes(group = 1), color = "red") +
  facet_wrap(~ Variable, scales = "free_x", nrow = 1,
             labeller = as_labeller(c(
               "Field_of_Study" = "Field of Study",
               "HS_GPA_Level" = "High School GPA",
               "Internships_Completed" = "Internships Completed",
               "Networking_Score" = "Networking Skill",
               "Projects_Completed" = "Uni Projects Completed"
             ))) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = NULL,
    y = "Average Number of Job Offers"
  )
```
:::
:::::

-   These two plot shows the distribution of **job offers** from **top 5** important predictors.
-   Both plots show that the **median** number of **job offers** falls between 2 and 3 for most categories.
-   Strongest positive signals come from **High-School GPA** and **Internships** because they show a stable increase in average offers as they grow.
-   It is also interesting to see **Art students** are more likely to receive **job offers** compare to **computer science students**.

---

## Conclusion

::: {.fragment}
- Practical experience and academic performance drive job offers.
:::

::: {.fragment}
- Internships and high school GPA are strong predictors.
:::

::: {.fragment}
- Field of study, networking, and projects matter—if approached strategically.
:::

::: {.fragment}
- Institutional prestige and soft skills had limited impact.
:::


---

## Recommendations

::: {.fragment}
- **Students** must focus on internships, GPA, and **balanced involvement** in co-curriculars.
:::

::: {.fragment}
- **Universities** should embed industry-linked experiences and align programs with job market needs.
:::

::: {.fragment}
- **Career Services** must promote strategic engagement beyond academics.
:::

