---
title: "slide"
format: 
  revealjs:
    code-fold: true
    echo: true
    toc: true
    multiplex: false
    show-slide-number: all
    show-notes: false
    controls: true
    width: 1280
    height: 750
embed-resources: true
---

```{r}
library(knitr)
library(tidyverse)
library(randomForest)
library(tidytext)
```

## page 3

------------------------------------------------------------------------

## page 4

```{r}
data = read.csv("data/education_career_success.csv")
```

------------------------------------------------------------------------

## page 5

------------------------------------------------------------------------

## Elements that contribute to receiving job offers {.smaller}

::::: columns
::: {.column width="60%"}
```{r, echo=FALSE}
outcome_var <- "Job_Offers"

predictors <- data %>%
  dplyr::select(-Student_ID, -Career_Satisfaction, -Work_Life_Balance) %>%
  dplyr::select(-all_of(outcome_var)) %>%
  names()

form <- as.formula(paste(outcome_var, "~", paste(predictors, collapse = " + ")))

set.seed(3)
rf_model <- randomForest(form, data = data, importance = TRUE)
rf_imp_df <- importance(rf_model, type = 1) %>%
  as.data.frame() %>%
  rownames_to_column("Variable") %>%
  rename(Importance = `%IncMSE`) %>%  
  arrange(desc(across(where(is.numeric))))

ggplot(rf_imp_df, aes(x = reorder(Variable, Importance), y = Importance)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Predictor",
    y = "% Increase in Accuracy"
  )
```
:::

::: {.column width="40%"}
-   This plot shows the importance of each predictor variable in the **random forest model** by using permutation.
-   The result shows that most important factors are related to your **academic performance** and **practical experience**.
-   It is also interesting to noted that **networking skill** rank much higher than **soft skills**.
:::
:::::

------------------------------------------------------------------------

## Association between important factors and job offers {.smaller}

::::: columns
::: {.column width="50%"}
```{r, echo=FALSE}
data <- data %>%
  mutate(HS_GPA_Level = cut(
    High_School_GPA,
    breaks = c(0, 0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4),
    labels = c( "0-0.5", "0.5-1", "1-1.5", "1.5-2", "2-2.5", "2.5-3", "3-3.5", "3.5-4"),
    right = TRUE,
    ordered_result = TRUE
  ))

data_plot <- data %>%
  select(Job_Offers, Internships_Completed, HS_GPA_Level, Field_of_Study, Networking_Score, Projects_Completed) %>%
  mutate(across(-Job_Offers, as.character)) %>%    
  pivot_longer(-Job_Offers, names_to = "Variable", values_to = "Category")


gpa_levels <- c("1.5-2", "2-2.5", "2.5-3", "3-3.5", "3.5-4")

data_plot_sorted <- data_plot %>%
  group_by(Variable) %>%
  group_modify(~{
    facet_df <- .x               # rows for this facet
    facet_nm <- .y$Variable[1]   # facet name (group key)

    if (facet_nm == "HS_GPA_Level") {
      # apply manual order
      facet_df %>%
        mutate(Category = factor(Category, levels = gpa_levels))
    } else {
      # automatic rule: numeric facets in ascending order,
      # text facets keep original order
      Category_num <- suppressWarnings(as.numeric(facet_df$Category))

      level_order <- if (all(is.na(Category_num))) {
        unique(facet_df$Category)                             # text → keep entry order
      } else {
        unique(facet_df$Category[order(Category_num, na.last = NA)])  # numbers → asc
      }

      facet_df %>%
        mutate(Category = factor(Category, levels = level_order))
    }
  }) %>%
  ungroup()
data_plot_sorted %>%
  ggplot(aes(x = Category, y = Job_Offers)) +
  geom_boxplot() +
  facet_wrap(~ Variable, scales = "free_x", nrow = 1,
             labeller = as_labeller(c(
                "Field_of_Study" = "Field of Study",
                "HS_GPA_Level" = "High School GPA",
                "Internships_Completed" = "Internships Completed",
                "Networking_Score" = "Networking Skill",
                "Projects_Completed" = "Uni Projects Completed"
              ))) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = NULL, y = "Number of Job Offers") 
```
:::

::: {.column width="50%"}
```{r, echo=FALSE}
data_plot_sorted %>%
  group_by(Variable, Category) %>%
  summarise(avg_offer = mean(Job_Offers), .groups = "drop") %>%
  group_by(Variable) %>%
  mutate(
    line_ok = Variable != "Field_of_Study",
    Category = {
      raw_cat <- as.character(Category)
      if (first(Variable) == "HS_GPA_Level") {
        left_val <- as.numeric(sub("-.*", "", raw_cat))
        factor(Category, levels = raw_cat[order(left_val)])
      } else if (all(grepl("^[0-9.]+$", raw_cat))) {
        factor(Category, levels = raw_cat[order(as.numeric(raw_cat))])
      } else if (all(grepl("^[0-9.]+-[0-9.]+$", raw_cat))) {
        right_val <- as.numeric(sub(".*-", "", raw_cat))
        factor(Category, levels = raw_cat[order(-right_val)])
      } else {
        factor(Category, levels = raw_cat[order(avg_offer, decreasing = TRUE)])
      }
    }
  ) %>%
  ungroup() %>%
  ggplot(aes(x = Category, y = avg_offer, group = Variable)) +
  geom_point(color = "steelblue", size = 2.5) +
  geom_line(data = function(df) dplyr::filter(df, line_ok), aes(group = 1), color = "red") +
  facet_wrap(~ Variable, scales = "free_x", nrow = 1,
             labeller = as_labeller(c(
               "Field_of_Study" = "Field of Study",
               "HS_GPA_Level" = "High School GPA",
               "Internships_Completed" = "Internships Completed",
               "Networking_Score" = "Networking Skill",
               "Projects_Completed" = "Uni Projects Completed"
             ))) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = NULL,
    y = "Average Number of Job Offers"
  )
```
:::
:::::

-   These two plot shows the distribution of **job offers** from **top 5** important predictors.
-   Both plots show that the **median** number of **job offers** falls between 2 and 3 for most categories.
-   Strongest positive signals come from **High-School GPA** and **Internships** because they show a stable increase in average offers as they grow.
-   It is also interesting to see **Art students** are more likely to receive **job offers** compare to **computer science students**.

------------------------------------------------------------------------

## page 8

------------------------------------------------------------------------

## page 9

------------------------------------------------------------------------

## page 10
